### 2020-11-19 paper/plot_macros.tex
partial refactor of tables

\newcommand{\commpcaplot}[1]{%
  \addplot[
  scatter, only marks,
  font=\tiny,
  scatter src=explicit symbolic,
  scatter/classes={
    general-interest={mark=*, blue},
    videogames={mark=*, red},
    technology={mark=*, green},
    sports={mark=*, cyan},
    female-focused={mark=*, magenta},
    other={mark=o, black}%
  },
  nodes near coords*={\community},
  visualization depends on={value \thisrow{community} \as \community}
  ] table [x=pca1, y=pca2, meta=category] {#1};
}

\newcommand{\commpcagroupplot}[2]{%
  \begin{groupplot}[
    group style={group size=2 by 1}, 
    yticklabels={,,}, xticklabels={,,},
    legend style={anchor=north, legend columns=1, font=\tiny},
    every node near coord/.append style={font=\sffamily\tiny}%
    ]
    \nextgroupplot
    \commpcaplot{#1}
    \nextgroupplot
    \commpcaplot{#2}
    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  \end{groupplot}
}

\newcommand{\confusionplot}[1]{%
  \addplot[
  matrix plot*,
  shader=faceted,
  mesh/cols=46,
  point meta=explicit,
  ] table [x=actual_comm, y=pred_comm, meta=#1] {floats/confusion.csv};
}

\newcommand{\confusiongroupplot}[2]{%
  \begin{groupplot}[
    group style={group size=2 by 1}, 
    enlargelimits=false,
    axis line style={draw=none},
    %axis equal,
    xmin=-1, xmax=46,
    ymin=-1, ymax=46,
    width=8cm,
    xticklabels={Advice,AskWomen,BabyBumps,CFB,Drugs,EDH,EarthPorn,Fantasy,GameDeals,GlobalOffensive,Jokes,Kappa,KerbalSpaceProgram,KotakuInAction,LifeProTips,MLS,MMA,MaddenUltimateTeam,TwoXChromosomes,Warframe,airsoft,bodybuilding,breakingmom,cars,cringe,eu4,exjw,explainlikeimfive,femalefashionadvice,food,heroesofthestorm,jailbreak,justneckbeardthings,oculus,pcmasterrace,photography,reddevils,relationships,rupaulsdragrace,stopdrinking,streetwear,techsupport,todayilearned,toronto,videos,xxfitness},
    yticklabels=none,
    xticklabel style={font=\sffamily\tiny, rotate=90},
    xtick=data,
    legend style={font=\sffamily\small},
    ]
    \nextgroupplot
    \confusionplot{#1}
    \nextgroupplot[colorbar right]
    \confusionplot{#2}
  \end{groupplot}
}

\newcommand{\pairwisecommsim}[1]{%
  \begin{tikzpicture}
    \begin{axis}[]
      \addplot[
        scatter, only marks,
        font=\tiny,
        scatter src=explicit symbolic,
        scatter/classes={
          general-interest={mark=*, blue},
          videogames={mark=*, red},
          technology={mark=*, green},
          sports={mark=*, cyan},
          female-focused={mark=*, magenta},
          other={mark=o, black},
          different={mark=x}
        },
        ] table [x=snap, y=#1, meta=meta] {floats/comm_sim.csv};
    \end{axis}
  \end{tikzpicture}
}

\newcommand{\pplscatter}[2]{%
  \begin{tikzpicture}
  \begin{axis}[]
  \addplot[
    only marks, black,
  ] table [x=#1, y=#2] {floats/cclm_lmcc_ppl.csv};
  \end{axis}
  \end{tikzpicture}
}