### 2021-05-03 paper/plot_macros.tex
updates to paper

\usepackage{lmodern}

\newcommand{\commscatter}[3][\community]{%
  \addplot[
  scatter, only marks,
  font=\tiny,
  scatter src=explicit symbolic,
  scatter/classes={
    general-interest={mark=*, blue},
    videogames={mark=*, red},
    technology={mark=*, green},
    sports={mark=*, cyan},
    female-focused={mark=*, magenta},
    other={mark=o, black}%
  },
  nodes near coords*={#1},
  nodes near coords align={vertical},
  every node near coord/.append style={font=\sffamily\tiny},
  visualization depends on={value \thisrow{community} \as \community},
  ] table [x={#2}, y={#3}, meta=category] {floats/comm.csv};
}

\newcommand{\confusionplot}[1]{%
  \addplot[
  matrix plot*,
  shader=faceted,
  mesh/cols=46,
  point meta=explicit,
  ] table [x=actual_comm, y=pred_comm, meta=#1] {floats/confusion.csv};
}

\newcommand{\confusionsingleplot}[1]{
  \begin{axis}[
    group style={group size=2 by 1}, 
    enlargelimits=false,
    axis line style={draw=none},
    %axis equal,
    xmin=-1, xmax=46,
    ymin=-1, ymax=46,
    width=8cm,
    xticklabels={Advice,AskWomen,BabyBumps,CFB,Drugs,EDH,EarthPorn,Fantasy,GameDeals,GlobalOffensive,Jokes,Kappa,KerbalSpaceProgram,KotakuInAction,LifeProTips,MLS,MMA,MaddenUltimateTeam,TwoXChromosomes,Warframe,airsoft,bodybuilding,breakingmom,cars,cringe,eu4,exjw,explainlikeimfive,femalefashionadvice,food,heroesofthestorm,jailbreak,justneckbeardthings,oculus,pcmasterrace,photography,reddevils,relationships,rupaulsdragrace,stopdrinking,streetwear,techsupport,todayilearned,toronto,videos,xxfitness},
    yticklabels=none,
    x tick label style={font=\sffamily\tiny,rotate=90,anchor=east},
    xtick=data,
    legend style={font=\sffamily\small},
    colorbar right,
    ]
    \confusionplot{#1}
  \end{axis}
}

\newcommand{\confusiongroupplot}[2]{%
  \begin{groupplot}[
    group style={group size=2 by 1}, 
    enlargelimits=false,
    axis line style={draw=none},
    %axis equal,
    xmin=-1, xmax=46,
    ymin=-1, ymax=46,
    width=8cm,
    xticklabels={Advice,AskWomen,BabyBumps,CFB,Drugs,EDH,EarthPorn,Fantasy,GameDeals,GlobalOffensive,Jokes,Kappa,KerbalSpaceProgram,KotakuInAction,LifeProTips,MLS,MMA,MaddenUltimateTeam,TwoXChromosomes,Warframe,airsoft,bodybuilding,breakingmom,cars,cringe,eu4,exjw,explainlikeimfive,femalefashionadvice,food,heroesofthestorm,jailbreak,justneckbeardthings,oculus,pcmasterrace,photography,reddevils,relationships,rupaulsdragrace,stopdrinking,streetwear,techsupport,todayilearned,toronto,videos,xxfitness},
    yticklabels=none,
    x tick label style={font=\sffamily\tiny,rotate=90,anchor=east},
    xtick=data,
    legend style={font=\sffamily\small},
    ]
    \nextgroupplot
    \confusionplot{#1}
    \nextgroupplot[colorbar right]
    \confusionplot{#2}
  \end{groupplot}
}

\newcommand{\pairwisecommsim}[2]{%
  \begin{tikzpicture}
    \begin{axis}[]
      \addplot[
        scatter, only marks,
        font=\tiny,
        scatter src=explicit symbolic,
        scatter/classes={
          general-interest={mark=*, blue},
          videogames={mark=*, red},
          technology={mark=*, green},
          sports={mark=*, cyan},
          female-focused={mark=*, magenta},
          other={mark=o, black},
          different={mark=x}
        },
        ] table [x=#1, y=#2, meta=category] {floats/comm_comm.csv};
    \end{axis}
  \end{tikzpicture}
}

\newcommand{\infogainboxplot}{
  \begin{groupplot}[
    group style={
      group size=1 by 2,
      every plot/.style={
        xmin=0.65,
        xmax=1.25,
      },
    },
    height=4cm, width=8cm,
    ytick={1,2,3,4,5,6,7,8},
    yticklabels={0, 1, 2, 3, 0, 1, 2, 3},
    ylabel={$l_c$},
    ]
  \nextgroupplot
  \addplot+[boxplot,draw=blue,mark=*,solid]
  table [y=lstm-3-0_info_gain] {floats/comm.csv};
  \addplot+[boxplot,draw=blue,mark=*,solid]
  table [y=lstm-3-1_info_gain] {floats/comm.csv};
  \addplot+[boxplot,draw=blue,mark=*,solid]
  table [y=lstm-3-2_info_gain] {floats/comm.csv};
  \addplot+[boxplot,draw=blue,mark=*,solid]
  table [y=lstm-3-3_info_gain] {floats/comm.csv};
  \nextgroupplot
  \addplot+[boxplot,draw=red,mark=*,solid]
  table [y=transformer-3-0_info_gain] {floats/comm.csv};
  \addplot+[boxplot,draw=red,mark=*,solid]
  table [y=transformer-3-1_info_gain] {floats/comm.csv};
  \addplot+[boxplot,draw=red,mark=*,solid]
  table [y=transformer-3-2_info_gain] {floats/comm.csv};
  \addplot+[boxplot,draw=red,mark=*,solid]
  table [y=transformer-3-3_info_gain] {floats/comm.csv};
  \end{groupplot}
}

% https://tex.stackexchange.com/questions/183778/how-can-several-boxplots-be-combined-into-a-single-figure-as-groups/183856#183856
\newcommand{\pplbycommbox}{
  \begin{axis}
    \addplot+[boxplot={draw position=-0.2, box extend=0.2}, draw=blue] table [y=lstm-3-0_ig] {floats/ig_by_comm.csv};
    \addplot+[boxplot={draw position=0.2, box extend=0.2}, draw=red] table [y=transformer-3-0_ig] {floats/ig_by_comm.csv};
    \addplot+[boxplot={draw position=.8, box extend=0.2}, draw=blue] table [y=lstm-3-1_ig] {floats/ig_by_comm.csv};
    \addplot+[boxplot={draw position=1.2, box extend=0.2}, draw=red] table [y=transformer-3-1_ig] {floats/ig_by_comm.csv};
    \addplot+[boxplot={draw position=1.8, box extend=0.2}, draw=blue] table [y=lstm-3-2_ig] {floats/ig_by_comm.csv};
    \addplot+[boxplot={draw position=2.2, box extend=0.2}, draw=red] table [y=transformer-3-2_ig] {floats/ig_by_comm.csv};
    \addplot+[boxplot={draw position=2.8, box extend=0.2}, draw=blue] table [y=lstm-3-3_ig] {floats/ig_by_comm.csv};
    \addplot+[boxplot={draw position=3.2, box extend=0.2}, draw=red] table [y=transformer-3-3_ig] {floats/ig_by_comm.csv};
  \end{axis}
}